---
title: "Dashboard"
author: "Christakakis Panagiotis"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r global, include=FALSE}
library(tidyverse)
library(tidygraph)
library(cshapes)
library(leaflet)
library(sf)
library(units)
library(lwgeom)
library(rgeos)
library(rgdal)
library(ggplot2)
library(prioritizr)
library(tmap)
library(ggraph)
library(ggdag)
library(igraph)
library(stringr)
```

Graph 1
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------

```{r input-1}
selectInput("layout", label = "Select the Layout of the graph",
            choices = c('kk', 'fr', 'linear', 'circlepack'), selected = 'kk')

```

Column
-----------------------------------------------------------------------

### Directed graph of capitals and their distances. 
With the use of proximity matrix we plot a directed graph of capitals that their countries neighbor. Each node corresponds to a capital and edges width represent the distances between the capitals.

```{r page-1}

cmap.2019 <- cshp(date=as.Date("2019-1-1"))
validcmap.2019 <- st_make_valid(cmap.2019)
cmap.2019_wgs84 <- st_transform(cmap.2019, "EPSG:4326")
cmap.2019_wgs84 <- st_make_valid(cmap.2019_wgs84)

distance_matrix <- distmatrix(as.Date("2019-1-1"), "capdist")

# Rows and columns are named after capitals.
rownames(distance_matrix) <- cmap.2019_wgs84$capname
colnames(distance_matrix) <- cmap.2019_wgs84$capname

# Convert distance matrix into df.
distance_matrix <- as.data.frame(distance_matrix)

# Add a new column with the names of capitals.
distance_matrix["capitals"] <- cmap.2019$capname

# Calculate proximity matrix from a map and convert it into df.
prox_matrix <- proximity_matrix(validcmap.2019,0)
prox_matrix <- as.matrix(prox_matrix)
prox_matrix <- as.data.frame(prox_matrix)

# Rows and columns are named after capitals.
rownames(prox_matrix) <- cmap.2019_wgs84$capname
colnames(prox_matrix) <- cmap.2019_wgs84$capname

# Add a new column with the names of capitals in order to gather them.
prox_matrix["capitals"] <- cmap.2019$capname
all_possible_connections <- gather(prox_matrix, key = "cap_name", value = "connection", 1:174)

# Make changes in order to fit together the two dataframes and lastly obtain the three
# columns we truly want (From - To - Distance).
dist_prox_together <- gather(distance_matrix, key = "cap_name", value = "dist", 1:174)
dist_prox_together['distances'] <- all_possible_connections$connection
true_connections <- dist_prox_together %>% filter(dist_prox_together$distances == 1) %>% arrange(desc(dist))
true_connections_no_duplicates <- distinct(true_connections, dist, .keep_all = TRUE)
final_connections <- true_connections_no_duplicates[,1:3]

graph_routes <- tidygraph::as_tbl_graph(final_connections)
nodes <- cmap.2019$capnames

graph_routes <- graph_routes %>%
  tidygraph::activate(nodes) %>%
  mutate(title = str_to_title(name),
         label = str_replace_all(title, " ", "\n")
  )
```


```{r plot-1}
renderPlot({
  graph_routes %>%
    ggraph(layout = input$layout) +
    geom_edge_link(aes(width = dist)) +
    scale_edge_width(range=c(0.2, 5)) +
    geom_node_text(aes(label = label, color = name), size = 3.5, repel = TRUE, show.legend = FALSE) +
    geom_edge_diagonal(color = "grey", alpha = 0.7)
})
```

Graph 2
=====================================

Column
-----------------------------------------------------------------------

### The shortest path between Athens and Copenhagen, considering distance as weight between them.
```{r page-2, include=FALSE}
stations <- graph_routes %>%
  tidygraph::activate(nodes) %>%
  pull(title)

# Define starting and ending point
from_one <- which(stations == "Athens")
to_one <-  which(stations == "Copenhagen")

# Find shortest weighted path
shortest_weighted <- graph_routes %>%
  morph(to_shortest_path, from_one, to_one, weights = dist, mode = "all")

# Re-arrange the order that the edges are drawn in the plot
# ensuring that the route will be drawn at the top.
shortest_weighted %>%
  mutate(selected_node = TRUE) %>%
  unmorph()

shortest_weighted <- shortest_weighted %>%
  mutate(selected_node = TRUE) %>%
  activate(edges) %>%
  mutate(selected_edge = TRUE) %>%
  unmorph() 

shortest_weighted <- shortest_weighted %>%
  activate(nodes) %>%
  mutate(selected_node = ifelse(is.na(selected_node), 1, 2)) %>%
  activate(edges) %>%
  mutate(selected_edge = ifelse(is.na(selected_edge), 1, 2)) %>%
  arrange(selected_edge)
```

```{r plot-2}
renderPlot({
  
    # Plot the shortest path
  shortest_weighted %>%
    ggraph(layout = "kk") +
    geom_edge_diagonal(aes(alpha = selected_edge), 
                       color = "black", show.legend = FALSE) +
    geom_node_text(aes(label = label, color =name, 
                       alpha = selected_node), size = 4.5, repel = TRUE, show.legend = FALSE) 
})
```
Graph 3
=====================================

Column
-----------------------------------------------------------------------

### The shortest path between Athens and Copenhagen, considering total number of nodes as weight between them.

```{r page-3, include=FALSE}
# Define starting and ending point
from_two <- which(stations == "Athens")
to_two <-  which(stations == "Copenhagen")

# Find shortest non-weighted path
shortest_non_weighted <- graph_routes %>%
  morph(to_shortest_path, from_two, to_two, mode = "all")

# Re-arrange the order that the edges are drawn in the plot
# ensuring that the route will be drawn at the top.
shortest_non_weighted %>%
  mutate(selected_node = TRUE) %>%
  unmorph()

shortest_non_weighted <- shortest_non_weighted %>%
  mutate(selected_node = TRUE) %>%
  activate(edges) %>%
  mutate(selected_edge = TRUE) %>%
  unmorph() 

shortest_non_weighted <- shortest_non_weighted %>%
  activate(nodes) %>%
  mutate(selected_node = ifelse(is.na(selected_node), 1, 2)) %>%
  activate(edges) %>%
  mutate(selected_edge = ifelse(is.na(selected_edge), 1, 2)) %>%
  arrange(selected_edge)

```

```{r plot-3}
renderPlot({
  # Plot the shortest path
  shortest_non_weighted %>%
    ggraph(layout = "kk") +
    geom_edge_diagonal(aes(alpha = selected_edge), 
                       color = "black", show.legend = FALSE) +
    geom_node_text(aes(label = label, color =name, 
                       alpha = selected_node), size = 4.5, repel = TRUE, show.legend = FALSE) 
})
```

Graph 4
=====================================
Column {.sidebar}
-----------------------------------------------------------------------

### Select for which capitals you want to find the shortest path.

```{r input-4}
selectInput("from", label = "From",
            choices = cmap.2019$capname, selected = 'Mbabane', multiple= FALSE)
selectInput("to", label = "To",
            choices = cmap.2019$capname, selected= 'Bangkok', multiple= FALSE)

```
Column
-----------------------------------------------------------------------

### The shortest path between two capitals considering distance as weight between them.

```{r page-4}

renderPlot({
  stations <- graph_routes %>%
  tidygraph::activate(nodes) %>%
  pull(title)
  
  from_two <- which(stations == toString(input$from))
  
  to_two <- which(stations == toString(input$to))
  
  # Find shortest non-weighted path
  shortest_non_weighted <- graph_routes %>%
                        morph(to_shortest_path, from_two, to_two, mode = "all")
  
  # Re-arrange the order that the edges are drawn in the plot
  # ensuring that the route will be drawn at the top.
  shortest_non_weighted %>%
    mutate(selected_node = TRUE) %>%
    unmorph()
  
  shortest_non_weighted <- shortest_non_weighted %>%
    mutate(selected_node = TRUE) %>%
    activate(edges) %>%
    mutate(selected_edge = TRUE) %>%
    unmorph() 
  
  shortest_non_weighted <- shortest_non_weighted %>%
    activate(nodes) %>%
    mutate(selected_node = ifelse(is.na(selected_node), 1, 2)) %>%
    activate(edges) %>%
    mutate(selected_edge = ifelse(is.na(selected_edge), 1, 2)) %>%
    arrange(selected_edge)
    
  # Plot the shortest path
  shortest_non_weighted %>%
    ggraph(layout = "kk") +
    geom_edge_diagonal(aes(alpha = selected_edge), 
                      color = "black", show.legend = FALSE) +
    geom_node_text(aes(label = label, color =name, 
                      alpha = selected_node), size = 6, repel = TRUE, show.legend = FALSE)
})
```
